Global adjoint tomography -- Model GLAD-M25 \\ Supplementary Material

### 2020-02-07 supplementary_material.tex
Updates from Overleaf

\title[Global Adjoint Tomography -- Model GLAD-M25]
  {Global adjoint tomography -- Model GLAD-M25\footnote{This paper is dedicated to Dimitri Komatitsch, computational geophysicist extraordinaire, who passed away unexpectedly shortly before its submission.}}

@@

%%---------------------------------------------
%
%   Summary
%
%%---------------------------------------------
\begin{summary}
Building on our first-generation global adjoint tomography model,
GLAD-M15~\citep{bozdaug2016global}, we present our second-generation transversely isotropic global model,
named GLAD-M25, which is the result of ten quasi-Newton tomographic iterations with an
earthquake database consisting of 1,480 events in the magnitude range $5.5\le M_w \le 7.2$,
an almost six-fold increase over the first-generation model.
We calculated fully 3D synthetic seismograms with a shortest period of 17~s based on a GPU-accelerated spectral-element wave
propagation solver which accommodates effects due to 3-D anelastic crust \& mantle structure, topography \& bathymetry, the ocean load, ellipticity, rotation, and self-gravitation.
We used an adjoint-state method to calculate Fr\'echet derivatives in 3D anelastic Earth models
facilitated by a parsimonious storage algorithm.
The simulations were performed on the Cray XK7 ``Titan'' and the IBM ``Summit'' at the Oak Ridge Leadership Computing Facility.
We quantitatively evaluated GLAD-M25 by assessing misfit reductions and traveltime \& amplitude anomaly histograms in twelve measurement
categories.
We performed similar assessments for a held-out data set consisting of 360 earthquakes,
with results comparable to the actual inversion.
We compare GLAD-M25 with several other global and regional models,
and highlight a variety of plumes and subduction zones.
\end{summary}

@@

\section{Introduction}

Construction of the first seismic tomographic models of the Earth dates back to the late 1970s~\citep{Aki77,Dziewonski77,SenTok77} and early 1980s~\citep[e.g.,][]{WD84,Nataf1984}.
Around the same time,
\cite{BaChLa77}, \cite{Lailly1983}, and~\cite{Tar84} formulated and adapted the theory of adjoint-state methods~\citep{Chavent1974} for exploration seismology with the goal of capturing the full physics of seismic wave propagation.
Mainly due to computational challenges,
it took until the late 2000s to see the first applications of adjoint-state methods in regional- and continental-scale earthquake seismology~\citep{tape2009adjoint,Fichtner09,zhu2012structure}.
The first global ``adjoint tomography'' model of the Earth's mantle,
GLAD-M15, was published in 2016~\citep{bozdaug2016global}, more than 30 years after the development of the original ``full waveform inversion'' (FWI) theory.

Current global seismic models are in general agreement regardless of data type and inversion strategy in terms of long-wavelength heterogeneity \citep{ritzwollerlavely1995,TW01,beckerboschi2002}.
However, discrepancies between models become noticeable as resolution increases.
Using accurate 3D simulations of seismic wave propagation and the computation of data sensitivities in 3D background models are key requirements for improving resolution in tomographic images on all scales~\citep{Tromp2020}.
Such detailed images are essential for understanding mantle dynamics and related surface tectonic processes
---for example the origin of hotspots and the forces behind plate motions and earthquakes.
Higher resolution wavespeed models are also important for accurately locating earthquakes, and are required from an engineering point of view to assess seismic hazard in earthquake prone regions and to detect nuclear explosions.

With these goals in mind,
in this study we use a GPU-accelerated version of the 3D spectral-element solver
SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02b} on the Cray supercomputer ``Titan'' and the IBM ``Summitt'' at the Oak Ridge Leadership Computing Facility (OLCF) for global adjoint tomography.
These simulations accommodate the full 3D complexity of global Earth models,
including 3D anelastic crust and mantle structure, self-gravitation, rotation, ellipticity, topography \& bathymetry, and the load of the oceans.
No compromises are made with regards to resolving the Earth's crust,
which is explicitly captured by the spectral-element mesh~\citep{tromp2010a},
thereby eschewing the need for ubiquitous ``crustal corrections''.
The ultimate goal is to use every single piece of information in seismograms,
a task made partly feasible based on the automated window selection tool FLEXWIN~\citep{maggi2009automated}.

Global adjoint tomography has a well-defined but complex workflow with multiple stages.
% including numerical simulations of synthetic seismograms (forward simulations) for all events,
% pre-processing of seismic data, making measurements, data assimilation, Fr\'echet derivative
% calculations (adjoint simulations), and post-processing of the gradient for quasi-Newton model
% updates.
It is essential to optimize, automate, and harden the entire process using workflow management tools, especially with large data sets.
For each iteration,
1,480 forward and adjoint simulations need to be performed,
generating a few Petabytes of wavefield files for the parsimonious-storage kernel calculation algorithm of~\cite{KoXiBoPeSaLiTr16}, and consuming 16~million OLCF node hours.
Accommodating inevitable hardware failures during this process is critical
to prevent contamination of the inversion results.
For these reasons, we are in the process of using EnTK as our
workflow management engine~\citep{EnTK2017}.
This workflow engine can automatically detected job failures,
both from the high-performance computing (HPC) system and via user-defined functions.
This facilitates tracking of tasks and semi-automatic job resubmission if necessary.

In this article we compare the results of our inversion with a number of previous tomographic models.
On a global scale, we present comparisons with model S362ANI$+$M~\citep{moulik2014anisotropic},
which is an updated version of model S362ANI~\citep{kustowski2008anisotropic},
the starting model for the GLAD-M15 inversion.
We also consider global shear wavespeed models S40RTS~\citep{ritsema2011s40rts},
TX2015~\citep{TX2015}, SEMUCB-WM1~\citep{french2015broad},
and SL2013sv~\citep{SchaefferLebedev13},
as well as compressional wavespeed models LLNL-G3Dv3 \citep{simmons2012llnl}, GAP-P4~\citep{fukao2013subducted}, and UU-P07~\citep{van2018atlas}.
On the scale of North America
we make comparisons with models US22~\citep{zhu2017radial},
a radially anisotropic model based on adjoint tomography using USArray data from 180 regional earthquakes,
and SL2013NA~\citep{schaeffer2014imaging},
a modification of global model SL2013sv focused on North America.
In Europe we use regional model EU60~\citep{zhu2015seismic} for comparison,
which is based on the assimilation of data from 190 regional earthquakes recorded by more than 700 European seismographic stations.
Finally, in Asia we use regional adjoint tomography model EARA2014~\citep{chen2015multiparameter} as a reference, a model based on data from 227~earthquakes recorded by more than 1,800~stations.

This article is organized as follows.
Sections~\ref{section:start}--\ref{section:data} describe the starting model,
earthquake database, and seismographic data set. 
Sections~\ref{section:misfit} and~\ref{section:parameterization} describe the misfit function minimized during the inversion process
and the related model parameterization.
In Section~\ref{section:workflow} we describe the adjoint tomography workflow in some detail.
Section~\ref{section:misfit_evolution} describes the evolution of the misfit, and in Section~\ref{section:evaluation} we evaluate the model based on a held-out data set of 360 earthquakes.
Finally, in Section~\ref{section:model}, we describe the model in detail,
offering many comparisons with various global and regional models, as described in the proceeding paragraph.
We conclude with a discussion of future opportunities and directions.

@@

As a continuation of our previous work~\citep{bozdaug2016global},
we used the first-generation model GLAD-M15 as our
starting model.
GLAD-M15 is a 3D transversely isotropic earth earth model, which combined
3D mantle model S362ANI~\citep{kustowski2008anisotropic}
with 3D crustal model CRUST2.0~\citep{bassin2000current} as its starting model.
In this study,
we continue to use the same transversely isotropic model parametrization.
Instead of relying on crustal corrections,
the mesh implementation of the crust in the spectral-element solver SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02b,PeKoLuMaLeCaLeMaLiBlNiBaTr11} enables us to accurately accommodate topography and bathymetry as well as variations in the Moho.
GLAD-M15 was constructed using a global database of 253 earthquakes.
The first 12 iterations of the GLAD-M15 inversion used three-component seismograms with a shortest period of 27~s,
and the final 3 iterations reduced this further to 17~s.
In this study we also used three-component seismograms with a shortest period of 17~s.

\section{Earthquakes}
\label{section:earthquakes}

@@

In addition to the 253 earthquakes used for the construction of GLAD-M15, we initially carefully selected 784 additional earthquakes
from the global Centroid-Moment Tensor (CMT) catalog~\citep[e.g.,][]{ekstrom2012global},
resulting in a total number of 1,040 earthquakes.
To ensure a good signal-to-noise ratio on a global scale,
the smallest moment magnitude in the database is set to~5.5,
thereby capturing more deep events,
and, to avoid complications associated with source dimension and directivity,
the largest moment magnitude is set to~7.2.

Before using the 1,040 earthquakes in our structural inversion,
we performed CMT inversions using model GLAD-M15,
the results of which are summarized in Fig.~\ref{fig:source_correction}.
To ensure even global coverage,
we used seismographic stations from the  Global Seismic Network (II, IU, IC, US, CU, and GT),
GEOFON~(GE), GEOSCOPE~(G), and several regional networks, such as MedNet~(MN),
the Brazilian Lithospheric Seismic Project~(BL), the Chilean National Seismic Network~(C),
and the Japan Meteorological Agency Seismic Network~(JP) (Fig.~\ref{fig:source_correction}b).
The number of stations used for CMT inversion usually ranges
from~150 to~500.

To determine the source parameters in the starting model,
we used the CMT inversion algorithm of \cite{liu2004spectral}.
This algorithm combines a normalized waveform difference misfit with an envelope difference misfit.
Specifically, the algorithm minimizes the misfit function
\begin{equation}
   \begin{split}
      \Phi =  \sum\limits_{c=1}^{C} \omega_c \sum\limits_{r=1}^{R_c} \omega_{cr}
       \sum\limits_{w=1}^{W_{cr}} \omega_{crw}\,
          & \left\{ \lambda\, \frac
              { \int \big[ d_w(t) - s_w(t - \Delta t) \big]^2 \mathrm{d}t}
              {\int \big[ d_w(t) \big]^2  \mathrm{d}t} \right.
       \\ & \quad \left. \mbox{} + (1 - \lambda)\, \frac
              {\int \big[ e(d_w(t)) - e(s_w(t - \Delta t)) \big]^2 \mathrm{d}t}
              {\int \big[ e(d_w(t)) \big]^2\mathrm{d}t} \right\}
              \quad ,
   \end{split}
\end{equation}
where~$d_w(t)$ denotes data in time window~$w$\,,
and~$s_w(t - \Delta t)$ the corresponding synthetic for model~GLAD-M15.
We allow the synthetics to shift relative to the data by an amount~$\Delta t$
--- effectively a station correction --- determined by cross correlation.
The envelope function is denoted by~$e(\,\cdot\,)$\,,
and the parameter~$\lambda$ determines the balance between fitting waveforms versus fitting envelopes.
The number of measurement categories is denoted by~$C$\,.
Following~\cite{ekstrom2012global},
seismograms were filtered between 50~s and 100~s
to select three-component body-wave windows,
and between 60~s and 100~s to select three-component
surface-wave windows,
resulting in six measurement categories, $c=1,\ldots,C$\,.
To balance their contributions,
each category is weighted by the reciprocal of the number of measurements in that
category, $\omega_c$\,.
The number of windows for a given receiver~$r$ in a given measurement category~$c$ is~$W_{cr}$\,.
Each such window is weighted equally, i.e.,~$\omega_{crw}=1$\,.
The number of receivers that records data in category~$c$ is denoted by~$R_c$\,.
As shown in Fig.~\ref{fig:source_correction}b,
the receivers are unevenly distributed across the globe,
with several dense arrays in the Northern Hemisphere and poor coverage in the Southern Hemisphere.
To balance the coverage,
following a strategy articulated by~\cite{Ruanetal2018},
we assign receiver weights~$\omega_{cr}$ based on the expression
\begin{equation}
\omega_{cr}^{-1} = N_c\,\sum_{r'=1}^{R_c} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_c}\right)^2\right]
\quad ,
\label{eq:spatial_weights}
\end{equation}
where~$N_c$ is a normalization factor for category~$c$,
and where~$\Delta_{rr'}$ denotes the angular distance between receivers~$r$ and~$r'$\,.
The reference angular distance~$\Delta_c$ for each category~$c$ needs to be chosen such that the condition 
number of the diagonal weighting matrix defined by Eqn.~(\ref{eq:spatial_weights}) is not too large.
The calculation of the weights may be abstracted as: given a distribution of
points on the unit sphere, determine a spatial weighting associated with each point.

The algorithm inverts for the six elements of the moment tensor
and the centroid location~(latitude, longitude, and depth).
These source inversions are computationally very expensive,
because for each of the 1,040~events nine full 3D forward simulations are required to obtain the source parameter Fr\'echet derivatives.

@@

Most earthquakes show a shallower depth after inversion
(Fig.~\ref{fig:source_correction}c),
consistent with our previous experiences~\citep[e.g.,][]{zhu2015seismic,chen2015multiparameter,bozdaug2016global} and with experiments by~\cite{hjorleifsdottir2010effects}.
We observed an average depth change of~$-2.62\pm2.49$~km relative to the global CMT solutions,
an average scalar moment change of~$5.31\pm3.91$\%,
and an average centroid time shift of~$-0.60\pm1.17$~s.
These are relatively minor corrections, especially in view of the significant expense of the source inversions.
For this reason, when we added another 440 earthquakes during iteration~22,
thereby bringing the total to~1,480 events,
we only performed a grid search to calibrate the centroid times and scalar moments.
Fig.~\ref{fig:event_1480} summarizes the characteristics of all 1,480 earthquakes used in this study.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/events_1480.pdf}
  \caption{\small{The 1,480~earthquakes used in this study. (a) Distribution of earthquakes. The color of each beach ball reflects its depth range, where blue designates events shallower than 50~km, green events between 50~km and 300~km, and red events deeper than 300~km. (b) \& (c) Histograms of earthquake moment magnitudes and depths.}}
  \label{fig:event_1480}
\end{figure}

\section{Seismic data}
\label{section:data}

The seismographic stations used in this study were carefully selected to ensure global coverage
and high data quality (Fig.~\ref{fig:stations}).
In addition to the seismographic  stations used for the source inversions,
we included all available data for our 1,480 event earthquake database from many data centers,
including IRIS, ORFEUS, INGV, IPGP, ETH, and GEONET.
Regional and temporary networks,
such as US Array~(TA),
Africa Array~(AF), the Canadian National Seismograph Network~(CN), Geoscience Australia~(AU),
the Antarctic Seismographic Argentinean Italian Network~(AI),
and the New Zealand National Seismograph Network~(NZ),
constitute a significant part
of our database and greatly improve coverage in certain regions.

\begin{figure}
  \includegraphics[width=0.8\textwidth]{figures/station_map.pdf}
  \caption{\small{Distribution of the 11,800~seismographic stations used in this study. Colors denote the number of events for which a given station contributes waveforms to the structural inversion. Stations with a number of event responses $<400$ are plotted in smaller size; these are usually temporary arrays deployed over a short period of time, frequently ocean bottom seismometers. The maximum number of event responses comes from ANMO, in Albuquerque, New Mexico, which contributed to 1,442 out of 1,480 earthquakes in the data set.}}
  \label{fig:stations}
  \centering
\end{figure}

\subsection{Adaptable Seismic Data Format}
\label{section:ASDF}

Conventional seismic data formats, such as SAC, involve one file per time series
plus related files with instrument response and event information.
Since every earthquake in the database is typically recorded by thousands of
instruments, I/O during the preprocessing stage of the adjoint tomography workflow
quickly cripples the file system. The Adaptable Seismic Data Format
(ASDF)~\citep[][ \texttt{https://seismic-data.org/}]{krischer2016adaptable}
was developed with complete reproducibility and fast parallel processing in mind.

In this context, there are four key issues ASDF resolves.
\begin{itemize}
    \item Robustness and stability: the data container is developed and maintained to ensure accuracy of scientific results.
    \item Data organization: the container is self-describing. Data, including waveform, source, and station information, are organized into well-defined structures.
    \item Reproducibility: the container enables scientists to keep track of which operations have been performed on the data so that results can be reproduced.
    \item Efficiency: the format provides easy mechanisms for parallel processing.
\end{itemize}

ASDF serves as a self-contained-and-explained data container while taking
full advantage of parallel computing. In our workflow,
one ASDF file contains all the information needed for processing one event,
including the seismic traces, an event information file (in QuakeML format), and station response files (in StationXML format).
The related Application Programming Interfaces (API)s are carefully designed for easy data extraction and parallel processing.

\section{Misfit function}
\label{section:misfit}

The misfit function to be minimized during the iterative inversion process must
be carefully constructed.
In this study,
it typically involves cross-correlation traveltimes
for body waves and frequency-dependent multitaper phase measurements for surface waves.
Measurements are made in several passbands of three-component seismograms rotated
into vertical, radial, and transverse components, resulting in a number of measurement
categories.
In this study, we consider four passbands, namely, a 17--40~s passband targeting body waves,
two 40--100~s passbands separately targeting body and surface waves,
and a 90--250~s passband targeting longer-period surface waves.
Each passband involves measurements on all three components,
which results in a total of twelve measurement categories.
Although the inversion is designed to minimize the overall misfit,
we track the misfit reduction in each of the categories to ensure that
the model is improving the fit to the data roughly equally across the board.
One of the biggest challenges in the construction of the misfit function is the
highly uneven distribution of earthquakes --- discussed in Section~\ref{section:earthquakes} --- and seismographic stations,
which must be counterbalanced by geographically weighting the data.
This issue is discussed in detail by~\cite{Ruanetal2018};
in this section we present a brief synopsis.

With these considerations in mind,
the overall misfit, $\Phi$, is defined as follows:
\begin{equation}
\label{eq:misfit}
\Phi = \sum_{s}^{S} \omega_s \sum_{c}^{C} \omega_{c} \sum_{r}^{R_{sc}} \omega_{scr} \sum_{w}^{W_{scr}} \omega_{scrw}\, \chi_{scrw}
\quad ,
\end{equation}
where~$S$ denotes the number of sources, $C$ the number of categories,
$R_{sc}$ the number of receivers recording source~$s$ in category~$c$,
and~$W_{scr}$ the number of measurement windows for source~$s$, category~$c$,
and receiver~$r$.
The misfit for a specific source~$s$, category~$c$, receiver~$r$, and window~$w$ is
\begin{equation}
  \chi_{scrw} = \frac{1}{\Delta\omega}\int_{\omega_1}^{\omega_2} \Big( \frac {\Delta \tau_{scrw}} {\sigma_{scrw}} \Big)^2\, \mathrm{d}\omega
\quad ,
\end{equation}
where~$\Delta \tau_{scrw}$ denotes a multi-taper frequency-dependent phase measurement over the frequency interval~$\Delta\omega=\omega_2-\omega_1$\,,
with associated uncertainties~$\sigma_{scrw}$\,.
For body waves, the window misfit is often simply the cross-correlation traveltime anomaly,
$\Delta T_{scrw}$\,, weighted by its standard deviation, i.e., $\chi_{scrw}=(\Delta T_{scrw}/\sigma_{scrw})^2$\,.
As in the source inversions,
the window weights are equal, $\omega_{scrw}=1$\,,
and the category weight, $\omega_c$\,, is the reciprocal of the number of measurements in that
category, $1/C$.
Earthquakes mainly occur along plate boundaries,
and most seismographic  stations are confined to the continents,
which leads to a very uneven distribution of sources and stations,
as illustrated in Figs.~\ref{fig:event_1480} and~\ref{fig:stations}.
Therefore, weighting is crucial in global tomography to balance this uneven sampling
of Earth's interior.
Our source and receiver weighting strategy,
encapsulated by the weights~$\omega_s$ and $\omega_{scr}$\,,
is developed to compensate for uneven spatial sampling.
Following the same strategy as for the receiver weights in the source inversions,
we define source and receiver weights as
\begin{equation}
\omega_{s}^{-1} = N\,\sum_{s'=1}^{S} \exp\left[\mbox{}-\left(\frac{\Delta_{ss'}}{\Delta}\right)^2\right]
\quad ,
\label{eq:source_weights}
\end{equation}
and
\begin{equation}
\omega_{scr}^{-1} = N_{sc}\,\sum_{r'=1}^{R_{sc}} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_{sc}}\right)^2\right]
\quad ,
\label{eq:receiver_weights}
\end{equation}
respectively.
Here~$N$ and~$N_{sc}$ are normalization factors,
and~$\Delta_{ss'}$ and~$\Delta_{rr'}$ denote angular distances between source and receiver pairs~$\{s,s'\}$ and~$\{r,r'\}$\,.
The reference angular distances~$\Delta$ and~$\Delta_{sc}$ need to be chosen based on the condition 
numbers of the diagonal weighting matrices defined by Eqns.~(\ref{eq:source_weights}) and~(\ref{eq:receiver_weights}).

\section{Model parameterization}
\label{section:parameterization}

We use the same transversely isotropic model parametrization as starting model GLAD-M15.
Such a model is described by the five Love parameters $A$, $C$, $L$, $N$, and $F$~\citep{Love27},
or, alternatively, using the mass density~$\rho$, in terms of the wavespeeds~$\alpha_v=\sqrt{C/\rho}$, $\alpha_h=\sqrt{A/\rho}$, $\beta_v=\sqrt{L/\rho}$, $\beta_h=\sqrt{N/\rho}$ and the dimensionless parameter $\eta=F/(A-2L)$~\citep{PREM,DT98}.
Assuming the radial anisotropy is due to shear anisotropy, these five parameters
may be further reduced to four by introducing the bulk sound speed,
$c=\sqrt{\kappa/\rho}$\,.
Therefore, the final four parameters are $c$, $\beta_v$, $\beta_h$, and $\eta$.
Transverse isotropy is confined to the upper mantle; the lower mantle is assumed to be isotropic.

Since density is difficult to constrain with seismic data,
density perturbations are scaled to isotropic (Voigt averaged) shear wavespeed perturbations based on the relationship~$\delta\ln\rho = 0.33\,\delta\ln\beta$~\citep{montagner1989petrological}.

Based on this parametrization,
the variation in the misfit function~(\ref{eq:misfit}) may be expressed as~\citep{zhu2015seismic,bozdaug2016global}
\begin{equation}
    \delta \Phi = \int_V
      (\delta \ln c\,K_c + \delta \ln\beta_v\,K_{\beta_v} + \delta \ln\beta_h\,K_{\beta_h} +
      \delta\ln\eta\,K_\eta) \mathrm{d}V
      \quad ,
\end{equation}
where~$K_c$, $K_{\beta_v}$, $K_{\beta_h}$, and $K_\eta$ denote the four Fr\'echet derivatives,
which are calculated based on an adjoint-state method~\citep[e.g.,][]{Plessix_2006_RAS,Tromp2005}.

@@

The adjoint tomography workflow is depicted in Fig.~\ref{fig:adjoint_workflow}.
It starts with the selection of earthquakes, as discussed in Section~\ref{section:earthquakes}.
Given the earthquake selection,
observed seismographic data and related response information are acquired,
as discussed in Section~\ref{section:data}.
For a given earthquake,
the corresponding seismograms and related response information are combined in a single ASDF file,
described in Section~\ref{section:ASDF}.
For a given earthquake data set, the conversion to ASDF needs to be performed once and for all.
In the following sections we highlight several important aspects of the workflow.

@@

The observed seismograms need to be compared to corresponding synthetic seismograms
for the purpose of making measurements.
The forward simulations are based on the global spectral-element solver
SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02a}, and take a CMT solution, a relevant list
of stations, and the current earth model as input.
The solver is GPU accelerated, and the simulations are performed on the Cray supercomputer ``Titan''
at the OLCF.
To simulate 120-min seismograms with a shortest period of 17~s on 384 NVIDIA
Tesla K20X GPUs takes approximately 10~min of wall clock time.

@@

The seismic data preprocessing workflow takes raw observed and synthetic data as input,
and generates adjoint sources for data assimilation as output.
From a computational perspective, the preprocessing workflow consumes only~1\%
of the overall computational requirements.
But from the perspective of the tomographic inversion, this is by far the most
important stage of the adjoint tomography workflow,
directly and fundamentally impacting the outcome.
Any ``bad data'' assimilated at this stage may contaminate the gradient and
ultimately the final model.

We use millions of seismograms generating tens of millions of measurement windows,
making data selection through human interaction impossible.
The preprocessing workflow automates this task,
most importantly via the time window selection tool FLEXWIN~\citep{maggi2009automated}.
Every seismogram and potential measurement goes through multiple checks and and
balances before acceptance or rejection.

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/window_colorbar_linear_Z.pdf}
  \caption{\small{Time-distance plot for vertical component 17--40~s body waves showing a measurement window density map.
  A total of 4,363,719 time windows were discretized into 1~degree epicentral distance and 1~s time bins.
  The maximum window count is 9,931 in the 85--86$^\circ$ distance and 1,375--1,376~sec time bin.
  Major traveltime branches are labeled. Note that the plot contains data from events with a wide variety of depths, so the branches are for identification purposes only.
  }}
  \label{fig:window_density_Z}
\end{figure}

>>>>

\title[Global Adjoint Tomography -- Model GLAD-M25 (Supplementary Material)]
  {Global adjoint tomography -- Model GLAD-M25 \\ Supplementary Material}

@@

%%---------------------------------------------
%
%   Summary
%
%%---------------------------------------------
\begin{summary}
This document contains supplementary material for the article entitled ``Global Adjoint Tomography -- Model GLAD-M25''.
\end{summary}

@@

@@

As a continuation of our previous work~\citep{bozdaug2016global},
we use model GLAD-M15 as our
starting model.
GLAD-M15 is a 3D transversely isotropic earth earth model, which combined
3D mantle model S362ANI~\citep{kustowski2008anisotropic}
with 3D crustal model CRUST2.0~\citep{bassin2000current} as its starting model.
We continue to use the same transversely isotropic model parametrization.
To accommodate attenuation,
we use radial anelastic model QL6~\citep{QL6}.

Instead of relying on crustal corrections,
the mesh implementation of the crust in the spectral-element solver SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02b,PeKoLuMaLeCaLeMaLiBlNiBaTr11} enables us to accurately accommodate topography and bathymetry as well as variations in the Moho.
The mantle and crust are inverted simultaneously.

GLAD-M15 was constructed using a global database of 253 earthquakes.
The first 12 iterations of the GLAD-M15 inversion used three-component seismograms with a shortest period of 27~s,
and the final 3 iterations reduced this further to 17~s.
In this study we also used three-component seismograms with a shortest period of 17~s.

\section{Source inversions}
\label{section:earthquakes}

@@

In addition to the 253 earthquakes used for the construction of GLAD-M15, we initially selected 784 additional earthquakes
from the global Centroid-Moment Tensor (CMT) catalog~\citep[e.g.,][]{ekstrom2012global},
resulting in a total number of 1,040 earthquakes.
To ensure a good signal-to-noise ratio on a global scale,
the smallest moment magnitude in the database is set to~5.5 (as in the global CMT catalog),
thereby capturing more deep events,
and, to avoid complications associated with source dimension and directivity,
the largest moment magnitude is set to~7.2.

Before using the 1,040 earthquakes in our structural inversion,
we performed CMT inversions using model GLAD-M15,
the results of which are summarized in Fig.~\ref{fig:source_correction}.
To optimize even global coverage,
we used seismographic stations from the  Global Seismic Network (II, IU, IC, US, CU, and GT),
GEOFON~(GE), GEOSCOPE~(G), and several regional networks, such as MedNet~(MN),
the Brazilian Lithospheric Seismic Project~(BL), the Chilean National Seismic Network~(C),
and the Japan Meteorological Agency Seismic Network~(JP) (Fig.~\ref{fig:source_correction}b).
The number of stations used for CMT inversion usually ranges
from~150 to~500.

To determine the source parameters in the starting model,
we used the CMT inversion algorithm of \cite{liu2004spectral}.
This algorithm combines a normalized waveform difference misfit with an envelope difference misfit.
Specifically, the algorithm minimizes the misfit function
\begin{equation}
   \begin{split}
      \Phi =  \sum\limits_{c=1}^{C} \omega_c \sum\limits_{r=1}^{R_c} \omega_{cr}
       \sum\limits_{w=1}^{W_{cr}} \omega_{crw}\,
          & \left\{ \lambda\, \frac
              { \int \big[ d_w(t) - s_w(t - \Delta t) \big]^2 \mathrm{d}t}
              {\int \big[ d_w(t) \big]^2  \mathrm{d}t} \right.
       \\ & \quad \left. \mbox{} + (1 - \lambda)\, \frac
              {\int \big[ e(d_w(t)) - e(s_w(t - \Delta t)) \big]^2 \mathrm{d}t}
              {\int \big[ e(d_w(t)) \big]^2\mathrm{d}t} \right\}
              \quad ,
   \end{split}
\end{equation}
where~$d_w(t)$ denotes data in time window~$w$\,,
and~$s_w(t - \Delta t)$ the corresponding synthetic for model~GLAD-M15.
We allow the synthetics to shift relative to the data by an amount~$\Delta t$
--- effectively a station correction --- determined by cross correlation.
The envelope function is denoted by~$e(\,\cdot\,)$\,,
and the parameter~$\lambda$ determines the balance between fitting waveforms versus fitting envelopes.
The number of measurement categories is denoted by~$C$\,.
Following~\cite{ekstrom2012global},
seismograms were filtered between 50~s and 100~s
to select three-component body-wave windows,
and between 60~s and 100~s to select three-component
surface-wave windows,
resulting in six measurement categories, $c=1,\ldots,C$\,.
To balance their contributions,
each category is weighted by the reciprocal of the number of measurements in that
category, $\omega_c$\,.
The number of windows for a given receiver~$r$ in a given measurement category~$c$ is~$W_{cr}$\,.
Each such window is weighted equally, i.e.,~$\omega_{crw}=1$\,.
The number of receivers that records data in category~$c$ is denoted by~$R_c$\,.
As shown in Fig.~\ref{fig:source_correction}b,
the receivers are unevenly distributed across the globe,
with several dense arrays in the northern hemisphere and poor coverage in the southern hemisphere.
To balance the coverage,
following a strategy articulated by~\cite{Ruanetal2018},
we assign receiver weights~$\omega_{cr}$ based on the expression
\begin{equation}
\omega_{cr}^{-1} = N_c\,\sum_{r'=1}^{R_c} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_c}\right)^2\right]
\quad ,
\label{eq:spatial_weights}
\end{equation}
where~$N_c$ is a normalization factor for category~$c$,
and where~$\Delta_{rr'}$ denotes the angular distance between receivers~$r$ and~$r'$\,.
The reference angular distance~$\Delta_c$ for each category~$c$ needs to be chosen such that the condition 
number of the diagonal weighting matrix defined by Eqn.~(\ref{eq:spatial_weights}) is not too large.
The calculation of the weights may be abstracted as: given a distribution of
points on the unit sphere, determine a spatial weighting associated with each point.

The algorithm inverts for the six elements of the moment tensor
and the centroid location~(latitude, longitude, and depth; duration is kept constant).
These source inversions are computationally very expensive,
because for each of the 1,040~events nine full 3D forward simulations are required to obtain the source parameter Fr\'echet derivatives.

@@

Most earthquakes are shallower after inversion
(Fig.~\ref{fig:source_correction}c),
consistent with our previous experiences~\citep[e.g.,][]{zhu2015seismic,chen2015multiparameter,bozdaug2016global} and with experiments by~\cite{hjorleifsdottir2010effects}.
We observed an average depth change of~$-2.62\pm2.49$~km relative to the global CMT solutions,
an average scalar moment change of~$5.31\pm3.91$\%,
and an average centroid time shift of~$-0.60\pm1.17$~s.
These are relatively minor corrections, especially in view of the significant expense of the source inversions.
For this reason, when we added another 440 earthquakes during iteration~22,
thereby bringing the total to~1,480 events,
we only performed a grid search to calibrate the centroid times and scalar moments.

% \subsection{Adaptable Seismic Data Format}
% \label{section:ASDF}

% Conventional seismic data formats, such as SAC, involve one file per time series
% plus related files with instrument response and event information.
% Since every earthquake in the database is typically recorded by thousands of
% instruments, I/O during the preprocessing stage of the adjoint tomography workflow
% quickly cripples the file system. The Adaptable Seismic Data Format
% (ASDF)~\citep[][ \texttt{https://seismic-data.org/}]{krischer2016adaptable}
% was developed with complete reproducibility and fast parallel processing in mind.

% In this context, there are four key issues ASDF resolves.
% \begin{itemize}
%     \item Robustness and stability: the data container is developed and maintained to ensure accuracy of scientific results.
%     \item Data organization: the container is self-describing. Data, including waveform, source, and station information, are organized into well-defined structures.
%     \item Reproducibility: the container enables scientists to keep track of which operations have been performed on the data so that results can be reproduced.
%     \item Efficiency: the format provides easy mechanisms for parallel processing.
% \end{itemize}

% ASDF serves as a self-contained-and-explained data container while taking
% full advantage of parallel computing. In our workflow,
% one ASDF file contains all the information needed for processing one event,
% including the seismic traces, an event information file (in QuakeML format), and station response files (in StationXML format).
% The related Application Programming Interfaces (API)s are carefully designed for easy data extraction and parallel processing.

@@

The adjoint tomography workflow is depicted in Fig.~\ref{fig:adjoint_workflow}.
It starts with the selection of earthquakes, as discussed in Section~\ref{section:earthquakes}.
Given the earthquake selection,
observed seismographic data and related response information are acquired.
For a given earthquake,
the corresponding seismograms and related response information are combined in a single ASDF file~\citep{krischer2016adaptable}.
For a given earthquake data set, the conversion to ASDF needs to be performed once and for all.
In the following sections we highlight several important aspects of the workflow.

@@

The forward simulations are based on the global spectral-element solver
SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02a}, and take a CMT solution, a list
of stations, and the current earth model as input.
The solver is GPU accelerated, and the simulations are performed on the Cray supercomputer ``Titan''
at the OLCF.
To simulate 120-min seismograms with a shortest period of 17~s on 384 NVIDIA
Tesla K20X GPUs takes approximately 10~min of wall clock time.

@@

The seismic data preprocessing workflow takes raw observed and synthetic data as input,
and generates adjoint sources for data assimilation as output.
From a computational perspective, the preprocessing workflow consumes only~1\%
of the overall computational requirements.
But from the perspective of the tomographic inversion, this is by far the most
important stage of the adjoint tomography workflow,
impacting the outcome.

We use millions of seismograms generating tens of millions of measurement windows,
making data selection through human interaction impossible.
The preprocessing workflow automates this task,
most importantly via the time window selection tool FLEXWIN~\citep{maggi2009automated}.

As illustrated in Fig.~\ref{fig:preprocess_workflow}, the preprocessing workflow
includes the following phases.
\begin{enumerate}
  \item Signal processing to remove the instrument response from observed data
    to recover ground displacement using ObsPy~\citep{obspy2010}. Both observed and synthetic data are bandpass
    filtered, and the horizontal components are rotated to obtain the radial and transverse components of motion.
  \item Window selection on a pair of
    observed and synthetic seismograms. Pyflex (a Python version
    of FLEXWIN by \citet{krischer2015}) is
    used to automatically generate windows where observed and
    synthetic data are sufficiently close to make measurements based on user defined
    criteria.
  \item Cross-correlation traveltime or multi-taper phase measurements in selected windows.
  \item Window sorting and cleaning based on statistical analyses of all measurements to eliminate outliers.
  \item Preliminary adjoint source construction based on the sorted windows.
  \item Calculation and assignment of source and receiver weights~\citep{Ruanetal2018}.
  \item Construction of the final, properly weighted, adjoint sources.
\end{enumerate}
